Towards Automated Merging of Code Clones in Object-Oriented Programming Languages\\- Work in Progress -

### 2019-05-13 sattose/paper.tex
Some small paper changes
Ana's latest feedback

\begin{abstract}
Duplication in source code can have a major negative impact on the maintainability of source code, as it creates explicit dependencies between fragments of code. Such explicit dependencies often cause bugs. In this study, we look into the opportunities to automatically refactor these duplication problems for object-oriented programming languages. To do this, we propose a method to detect clones that are suitable for refactoring. This method focuses on the context and scope of clones, ensuring our refactoring improves the design and does not leave side effects after it is applied.

@@

Refactoring is used to improve the quality-related attributes of a codebase (maintainability, performance, etc.) without changing the functionality. Many methods were introduced to aid the process of refactoring~\cite{fowler1999refactoring, wake2004refactoring}, and are integrated into most modern IDE's. However, most of these methods still require a manual assessment of where and when to apply them. This means refactoring is either a significant part of the development process~\cite{lientz1978characteristics, mens2004survey}, or does not happen at all~\cite{mens2003refactoring}. For a large part, proper refactoring requires domain knowledge. However, there are also refactoring opportunities that are rather trivial and repetitive to execute. In this research, we investigate the extent to which code clones can be automatically refactored.

A survey by Roy et al. \cite{roy2007survey} describes different types of clones. Most clone detection tools focus on finding clones corresponding with these definitions. In this paper, we outline challenges with these clone type definitions when considered in a refactoring context. We next propose solutions to these problems that would enable the detection of clones that should be refactored, rather than fragments of code that are just similar.

@@

We conduct our research on a large corpus of open source projects. We focus mainly on the Java programming language as refactoring opportunities feature paradigm and programming language dependent aspects~\cite{choi2011extracting}. However, most practices featured in this research will also be applicable to other object-oriented languages, like C\#.

In this research, we strive to improve upon the current state-of-the-art in clone refactoring \cite{fontana2015duplicated, alwaqfi2017refactoring} by building a clone refactoring tool that automatically applies refactorings to a large percentage of clones found. The design decisions for this tool are made on basis of data gathered from a large corpus of software systems.

@@

\subsection{Advantages of clone classes over clone pairs}\label{chap:cloneclasses}
Regarding clone detection, there is a lot of variability in literature whether clone pairs or clone classes should be considered for detection. In this study we focus on clone classes, because of the advantages for refactoring. Clone pairs do not provide a general overview of all entities containing the clones, with all their related issues and characteristics \cite{fontana2012duplicated}. Although clone classes are harder to manage, they provide all information needed to plan a suitable refactoring strategy, since this way all instances of a clone are considered. There is also another issue that results from grouping clones by pairs: clone reference amount increases according to the binomial coefficient formula (two clones form a pair, three clones form three pairs, four clones form six pairs, and so on), which causes a heavy information redundancy \cite{fontana2012duplicated}.

@@

\subsection{Related work in clone refactoring tools}
The Duplicated Code Refactoring Advisor (DCRA) looks into refactoring opportunities for clone pairs \cite{fontana2012duplicated, fontana2015duplicated}. This tool only focuses on refactoring clone pairs, with the authors arguing that \textit{clone pairs are much easier to manage when considered singularly.} As intermediate steps, the authors measure a corpus of Java systems for some clone-related properties of the systems, like the relation (in terms of inheritance) between code fragments in a clone pair. We further look into these measurements in section \ref{chap:relationsinstances}.

Aries \cite{higo2004aries, higo2008metric} focuses on the detection of refactorable clones. They do this based on the relation between clone instances through inheritance, similar to Fontana et al. \cite{fontana2012duplicated}, Aries also proposes a refactoring: if two clone classes are siblings of each other (share the same superclass), they propose to perform ``Extract method'' and ``Pull up method'' sequentially. This tool only proposes the refactoring, and does not provide help in the process of applying the refactoring.

There are various other researches that investigate refactoring code clones \cite{alwaqfi2017refactoring, chen2018clone, koni2001scenario}. However, all of these tools only support a subset of all harmful clones that are found. Also, these tools are limited to suggesting refactoring opportunities, rather than actually applying refactorings where suitable. Finally, all published approaches have limitations, such as false positives in their clone detection \cite{chen2018clone} or being limited to clone pairs \cite{higo2008metric}.

@@

\subsection{Survey on Clone Detection Tools}
\label{ch:tool-overview}
We conducted a short survey on (recent) clone detection tools that we could use to analyze refactoring possibilities. The results of our survey are displayed in table~\ref{table:dettools}. We chose a set of tools that are open source and can analyze either Java or Python. Next we formulate the following four criteria by which we analyze these tools:
\begin{enumerate}
    \item \textbf{Should find clones in any context.} Some tools only find clones in specific contexts, such as only method-level clones. We want to perform an analysis on all clones in projects to get a complete overview.
\item \textbf{Should find all clones in created test projects.} We have created a number of test projects that we used to assess the validity of clone detection tools. On basis of this, we checked whether clone detection tools can correctly find clones in diverse contexts.
\item \textbf{Can analyse resolved symbols.} When detecting clones for refactoring purposes, it is important that clone instances are actually equal. Sometimes, to verify actual equality (not just textual equality), it is required to consider resolved symbols (this is described more elaborately in section \ref{chap:type1clones}). Because of this, we want to use a clone detection tool that can analyze such structures.
\item \textbf{Extensive detection configuration.} We aim to exclude expressions/statements from matching (more about our rationale in section~\ref{chap:clonetypes}). To achieve this, the tool needs to be able to allow those threshold changes. This can be either through simple changes of the source code, or by using some configuration file.
\end{enumerate}

@@

None of the tools we considered implement all our criteria, so we decided to implement our own clone detection tool: CloneRefactor\footnote{CloneRefactor (WIP) is available on GitHub: \url{https://github.com/SimonBaars/CloneRefactor}. This repo contains all scripts that were used to retrieve the data that is displayed in this paper.}.

\subsection{CloneRefactor}
A 2016 survey by Gautam \cite{gautam2016various} focuses more on various techniques for clone detection. We decided to combine AST- and Graph-based approaches for clone detection, similar to Scorpio \cite{higo2013revisiting, kamalpriya2017enhancing}. However, instead of building a dependency graph, we build a similarity graph of statements linking to similar statements. The graph built by CloneRefactor is shown in Figure \ref{fig:clonerefactor}.

@@

We decided to base our tool on the JavaParser library \cite{tomassetti2017javaparser}, as it supports rewriting the AST back to Java code and is compatible with all modern Java versions (Java 1-12). We collect each statement and declaration and compare those to find duplicates. This way we build a graph of each statement/declaration linking to each subsequent statement/declaration (horizontally) and linking to each of its duplicates (vertically). On basis of this graph we detect clone classes. We challenge completed clone classes against the thresholds that are being used, and remove the clone classes that do not adhere to the thresholds.

\subsubsection{Thresholds}\label{chap:thresholds}
CloneRefactor works on basis of three thresholds for finding clones:
\begin{enumerate}
  \item \textbf{Number of statements/declarations:} The number of statements/declarations that should be equal/similar for it to be considered a clone.
  \item \textbf{Number of tokens:} The number of tokens (excluding whitespace, end-of-line terminators and comments) that should be equal/similar for it to be considered a clone.
  \item \textbf{Number of lines:} The minimum amount of lines (excluding lines that do not contain any tokens) that should be equal/similar for it to be considered a clone.
\end{enumerate}
Of course, if any threshold is set to zero, it will be ignored, thus not all thresholds have to be used at all times. We consider ``number of lines'' to be the least important, as it highly depends on the programmer of the codebase (and we want to prevent this dependence). On basis of manual assessment, we have determined that setting the ``number of statements/declarations'' to 6 ensures that most non-harmful clones are filtered out. On the downside, this also filters out some harmful antipatterns. For instance, if a cloned statement has many tokens we might want to consider it a clone even if it spans less than 6 statements (as a cloned line with many tokens is more harmful than one with few).

The measurements in our study use the following thresholds:
\begin{itemize}
  \item \textbf{Number of statements/declarations:} 3
  \item \textbf{Number of tokens:} 12
  \item \textbf{Number of lines:} 1
\end{itemize}

\section{Addressing problems with clone type definitions}\label{chap:clonetypes}
We propose solutions for shortcomings in type 1, type 2 and type 3 clones~\cite{roy2007survey} (see section \ref{chap:backgroundclonetypes} for these definitions) to increase the chance that we can merge the clone while improving the design. Due to the serious challenges involved in their detection and refactoring, type 4 clones are not considered in this study.

\subsection{Type 1R clones} \label{chap:type1clones}
Type 1 clones, as defined in literature~\cite{roy2007survey}, are identical clone fragments except for variations in whitespace and comments. However, when two clone fragments are textually identical, it does not yet indicate that they are actually identical. Even when textually equal, method calls can refer to different methods, type declarations can refer to different types and even variables can be of a different type. This could invalidate a refactoring opportunity for such a code fragment. Therefore, we propose an alternative definition of type 1 clones. We have named this it ``type 1R clones'', to indicate that this definition is separate from type 1 clones in literature. The ``R'' stands for refactoring, to indicate that this is type is refactoring oriented (and may be less suitable for, for instance, large scale clone analysis). Type 1R clones differ from type 1 clones by the following aspects:
\begin{itemize}
  \item \textbf{Compare the equality of the fully qualified method signature for method references.} If an identifier is fully qualified, it means we specify the full location of its declaration (e.g. \texttt{com.simonbaars.fruitgame.Apple} for an \texttt{Apple} object). This way we can validate whether two method references, like method calls, are actually equal. In the method signature, not only the fully qualified identifier of the method should be considered, but also the type of all its arguments. This way we can be sure that two potentially cloned method references do not point to overloaded variants (in a case that the data type of arguments is overloaded).
  \item \textbf{Compare the equality of the fully qualified identifier for type references.} This way we can be sure that two referenced types are actually equal, and that they are not just two types with the same name.
  \item \textbf{Compare the equality of the fully qualified identifier for variable usages.} Two cloned lines might use a variable with the same name, but different types. This might pose serious challenges on refactoring, as the variables might not concern the same object or primitive. To check this, we need to track the declaration of variables and from this infer the fully qualified identifier of its type.
  \item \textbf{Compare the equality of the fully qualified identifier for method references and call signatures.} This is to prevent a change in functionality after merging the clone, as methods with equal names/parameters can still contain different functionality.
\end{itemize}

\subsection{Type 2R clones}
By its definition in literature, type 2 clones allow any change in identifiers, literals, types, layout, and comments. For refactoring purposes, this definition is unsuitable. If we allow any change in identifiers, literals, and types, we cannot distinguish between different variables, different types and different method calls anymore. This could render two methods that have an entirely different functionality as clones. Merging such clones can be harmful instead of helpful.

@@

\begin{itemize}
  \item \textbf{Considering types:} The definition of type 2 clones states that types should not be considered. We disagree because types can make a significant change to the meaning of a code segment and thus whether this segment should be considered a clone.
  \item \textbf{Having a distinction between different variables:} By the definition of type 2 clones, any identifiers would not be taken into account. We agree that a difference in identifiers may still result in a harmful clone, but we should still consider the distinction between different variables. For instance, if we call a method like this: \texttt{myMethod(var1, var2)}, or call this method like this: \texttt{myMethod(var1, var1)}. Even if the variables have the same type, the distinction between the variables is important to ensure the functionality is the same after merging.
  \item \textbf{Defining a threshold for variability in literals:} By the definition of type 2 clones any literals would not be taken into account. We agree, as when merging the clone (for instance by extracting a method), we can simply turn the literal into a method parameter. However, we would argue that thresholds matter here. How many literals may differ for the segment still to be considered a clone with another segment? We need to define a threshold to be sure that, by merging, we are not replacing a code fragment by a worse maintainable design.
  \item \textbf{Consider method call signatures and define a threshold for variability in method calls:} As type-2 clones allow changes in identifiers, also the names of called methods may vary. However, because of this, completely different methods can be called in cloned fragments as a result. This poses serious challenges on refactoring and makes it more disputable whether such a clone is actually harmful. This is because different method identifiers can describe a completely different functionality. Therefore, we suggest considering the call signatures of cloned methods when they are compared. We can allow variability in the rest of method identifiers by passing the function as a parameter. To limit the amount of parameters required we also recommend defining a threshold for variability in method call expressions, so only a limited number of method calls can vary.
\end{itemize}

@@

Clone instances in type 3 clones are almost always different in functionality. As we have to ensure equal functionality after merging the clone, we have to wrap the difference in statements between the clone instances in conditional blocks (either if-statements or switch-statements). We can then pass a variable to indicate which path should be taken through the code (either a boolean or an enumeration). Such a refactoring would make added statements that are contiguous less harmful for the design than added statements that are scattered throughout the cloned fragment.

@@

In our CloneRefactor tool we use JavaParser \cite{tomassetti2017javaparser}. JavaParser has a build in symbol solver, which can automatically resolve types, method calls and variables. However, because of this, CloneRefactor does require all libraries that the software project requires (apart from the Java Standard Library). If these are not available, CloneRefactor will estimate types on basis of the information that is available.

@@

\begin{table}[H]
  \begin{center}
  \caption{CloneRefactor results for Java projects corpus \cite{githubCorpus2013}.} \label{table:general}
  \medskip
\begin{tabular}{|l|l|}
\hline
Amount of projects                                                                                      & 1,361      \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of lines (excluding\\whitespace, comments and newlines.)\end{tabular} & 1,414,996  \\ \hline
Amount of statements/declarations                                                                       & 1,212,189  \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of tokens (excluding\\whitespace, comments and newlines.)\end{tabular} & 11,643,194 \\ \hline
\end{tabular}
\end{center}
\end{table}

@@

\begin{table}[H]
  \begin{center}
  \caption{CloneRefactor clone detection results for the two different algorithms.} \label{table:clonedet}
  \medskip
\begin{tabular}{|l|l|l|}
\hline
Extra & \textbf{Type 1} & \textbf{Type 1R} \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of lines cloned\end{tabular} & 200,362 & 129,519 \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of statements/\\declarations cloned\end{tabular} & 182,466 & 118,980 \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of tokens cloned\end{tabular} & 1,582,845 & 973,596 \\ \hline
\end{tabular}
\end{center}
\end{table}

>>>>

\begin{abstract}
Duplication in source code can have a major negative impact on the maintainability of source code, as it creates explicit dependencies between fragments of code. Such explicit dependencies often cause bugs. In this study, we look into the opportunities to automatically refactor these duplication problems for object-oriented programming languages. We propose a method to detect clones that are suitable for refactoring. This method focuses on the context and scope of clones, ensuring our refactoring improves the design and does not created side effects.

@@

Refactoring is used to improve the quality-related attributes of a codebase (maintainability, performance, etc.) without changing the functionality. Many methods were introduced to aid the process of refactoring~\cite{fowler1999refactoring, wake2004refactoring}, and are integrated into most modern IDE's. However, most of these methods still require a manual assessment of where and when to apply them. This means refactoring is either a significant part of the development process~\cite{lientz1978characteristics, mens2004survey}, or does not happen at all~\cite{mens2003refactoring}. For a large part, proper refactoring requires domain knowledge. However, there are also refactoring opportunities that are rather trivial and repetitive to execute. Our goal is investigating to what extend code clones can be automatically refactored.

A survey by Roy et al. \cite{roy2007survey} describes various ways in which clones can be identified. Most clone detection tools focus on finding clones that align with these definitions. In this paper, we outline challenges with these clone type definitions when considered in a refactoring context. We next propose solutions to these problems that would enable the detection of clones that should be refactored, rather than fragments of code that are just similar.

@@

We conduct our research on a large corpus of open source projects. We focus mainly on the Java programming language as refactoring opportunities feature paradigm and programming language dependent aspects~\cite{choi2011extracting}. However, most practices featured currently in our work will also be applicable to other object-oriented languages, like C\#. This is because these programming languages share many similarities regarding refactoring opportunities.

Our end goal is to improve upon the current state-of-the-art in clone research \cite{fontana2015duplicated, alwaqfi2017refactoring} by building a clone refactoring tool that automatically applies refactorings to a large percentage of clones found. The design decisions for this tool are made on basis of data gathered from a large corpus of software systems together with our own experience and findings from literature.

@@

\subsection{Advantages of clone classes over clone pairs}\label{chap:cloneclasses}
Regarding clone detection, there is a lot of variability in literature whether clone pairs or clone classes should be considered for detection. We focus on clone classes, because of the advantages for refactoring. Clone pairs do not provide a general overview of all entities containing the clones, with all their related issues and characteristics \cite{fontana2012duplicated}. Although clone classes are harder to manage, they provide all information needed to plan a suitable refactoring strategy, since this way all instances of a clone are considered. Another issue that results from grouping clones by pairs: clone reference amount increases according to the binomial coefficient formula (two clones form a pair, three clones form three pairs, four clones form six pairs, and so on), which causes a heavy information redundancy \cite{fontana2012duplicated}.

@@

\subsection{Related work in clone refactoring tools}
The Duplicated Code Refactoring Advisor (DCRA) looks into refactoring opportunities for clone pairs \cite{fontana2012duplicated, fontana2015duplicated}. DCRA only focuses on refactoring clone pairs, with the authors arguing that \textit{clone pairs are much easier to manage when considered singularly.} As intermediate steps, the authors measure a corpus of Java systems for some clone-related properties of the systems, like the relation (in terms of inheritance) between code fragments in a clone pair. We further look into these measurements in section \ref{chap:relationsinstances}.

Aries \cite{higo2004aries, higo2008metric} focuses on the detection of refactorable clones. They do this based on the relation between clone instances through inheritance, similar to Fontana et al. \cite{fontana2012duplicated}. Aries also proposes a refactoring: if two clone classes are siblings of each other (share the same superclass), they propose to perform ``Extract method'' and ``Pull up method'' sequentially. This tool only proposes the refactoring, and does not provide help in the process of applying the refactoring.

We investigated several research efforts that look into code clone refactoring \cite{alwaqfi2017refactoring, chen2018clone, koni2001scenario}. However, all of these tools only support a subset of all harmful clones that are found. Also, these tools are limited to suggesting refactoring opportunities, rather than actually applying refactorings where suitable. Finally, all published approaches have limitations, such as false positives in their clone detection \cite{chen2018clone} or being limited to clone pairs \cite{higo2008metric}.

@@

\subsection{Survey on Clone Detection Tools}
\label{ch:tool-overview}
We conducted a short survey on (recent) clone detection tools that we could use to analyze refactoring possibilities. The results of our survey are displayed in table~\ref{table:dettools}. We chose a set of tools that are open source and can analyze a popular object-oriented programming language. Next, we formulate the following four criteria by which we analyze these tools:
\begin{enumerate}
    \item \textbf{Should find clones in any context.} Some tools only find clones in specific contexts, such as only method-level clones. We want to perform an analysis on all clones in projects to get a complete overview.
\item \textbf{Should find all clones in created control projects.} We assembled a number of test projects to assess the validity of clone detection tools. On basis of this, we checked whether clone detection tools can correctly find clones in diverse contexts.
\item \textbf{Can analyse resolved symbols.} When detecting clones for refactoring purposes, it is important that clone instances can be merged. Sometimes, textual equality between code fragments does not imply that these can be merged (this is described more elaborately in section \ref{chap:type1clones}). Because of this, we want to use a clone detection tool that can analyze such structures.
\item \textbf{Extensive detection configuration.} We aim to exclude expressions/statements from matching (more about our rationale in section~\ref{chap:clonetypes}). To achieve this, the tool needs to be able to allow those threshold changes. This can be either through simple changes of the source code, or by using some configuration file.
\end{enumerate}

@@

None of the state-of-the-art tools we identified implement all our criteria, so we decided to implement our own clone detection tool: CloneRefactor\footnote{CloneRefactor (WIP) is available on GitHub: \url{https://github.com/SimonBaars/CloneRefactor}. This repository contains all scripts that were used to retrieve the data that is displayed in this paper.}.

\subsection{CloneRefactor}
A 2016 survey by Gautam \cite{gautam2016various} focuses more on various techniques for clone detection. We decided to combine AST- and Graph-based approaches for clone detection, similar to Scorpio \cite{higo2013revisiting, kamalpriya2017enhancing}. However, instead of building a dependency graph, we build a similarity graph of statements (linking to similar statements). The graph built by CloneRefactor is shown in Figure \ref{fig:clonerefactor}.

@@

We decided to base our tool on the JavaParser library \cite{tomassetti2017javaparser}, as it supports rewriting the AST back to Java code and is compatible with all modern Java versions (Java 1-12). We collect each statement and declaration and compare those to find duplicates. We build a graph of each statement/declaration linking to each subsequent statement/declaration (horizontally) and linking to each of its duplicates (vertically). On basis of this graph we detect clone classes. We compare clone classes against the thresholds, and remove the clone classes that do not pass the test.

\subsubsection{Thresholds}\label{chap:thresholds}
CloneRefactor works on basis of three thresholds for finding clones:
\begin{enumerate}
  \item \textbf{Number of statements/declarations:} The number of statements/declarations that should be equal/similar for it to be considered a clone.
  \item \textbf{Number of tokens:} The number of tokens (excluding whitespace, end-of-line terminators and comments) that should be equal/similar for it to be considered a clone.
  \item \textbf{Number of lines:} The minimum amount of lines (excluding lines that do not contain any tokens) that should be equal/similar for it to be considered a clone.
\end{enumerate}
Of course, if any threshold is set to zero, it will be ignored, thus not all thresholds have to be used at all times. We consider ``number of lines'' to be the least important, as it highly depends on the developer of the codebase (and we want to prevent this dependence). Empirical research indicates that setting the ``number of statements/declarations'' to 3 ensures that many non-harmful clones are filtered out, as lower thresholds detect many clones that pose no harm to the design of the code. On the downside, this also filters out some harmful antipatterns. For instance, if a cloned statement has many tokens we might want to consider it a clone even if it spans a single statement (as a cloned line with many tokens is more harmful than one with few).

The measurements in our study use the following thresholds:
\begin{itemize}
  \item \textbf{Number of statements/declarations:} 1
  \item \textbf{Number of tokens:} 12
  \item \textbf{Number of lines:} 1
\end{itemize}

\section{Addressing problems with clone type definitions}\label{chap:clonetypes}
For each of type 1-3 clones~\cite{roy2007survey} we list our solutions to their shortcomings (see section \ref{chap:backgroundclonetypes} for these definitions) to increase the chance that we can merge the clone while improving the design. Due to the serious challenges involved in their detection and refactoring, type 4 clones are not considered in this study.

\subsection{Type 1R clones} \label{chap:type1clones}
Type 1 clones are identical clone fragments except for variations in whitespace and comments. However, when two clone fragments are textually identical, it does not yet indicate that they are functionally identical. Although textually equal, method calls can still refer to different methods, type declarations can still refer to different types and even variables can be of a different type. In such cases refactoring opportunities could be invalidated. Therefore, we propose an alternative definition of type 1 clones. We have named it ``type 1R clones''. The ``R'' stands for refactoring-oriented (and may be less suitable for other analyses). Type 1R clones differ from type 1 clones as follows:
\begin{itemize}
  \item \textbf{Compare the equality of the fully qualified method signature for method references.} If an identifier is fully qualified, it means we specify the full location of its declaration (e.g. \texttt{com.simonbaars.fruitgame.Apple} for an \texttt{Apple} object). This way we can validate whether two method references, like method calls, are functionally equal. In the method signature, not only the fully qualified identifier of the method should be considered, but also the type of all its arguments. This way we can be sure that two potentially cloned method references do not point to overloaded variants (in a case that the data type of arguments is overloaded).
  \item \textbf{Compare the equality of the fully qualified identifier for type references.} This way we can be sure that two referenced types are actually equal, and that they are not just two types with the same name.
  \item \textbf{Compare the equality of the fully qualified identifier for variable usages.} Two cloned lines might use a variable with the same name, but different types. This might pose serious challenges on refactoring, as the variables might not concern the same object or primitive. To check this, we need to track the declaration of variables and from this infer the fully qualified identifier of its type.
  \item \textbf{Compare the equality of the fully qualified identifier for method references and call signatures.} This is to prevent a change in functionality after merging the clone, as methods with equal names/parameters can still contain different functionality.
\end{itemize}

\subsection{Type 2R clones}
Type 2 clones allow any change in identifiers, literals, types, layout, and comments. For refactoring purposes, this definition is unsuitable; if we allow any change in identifiers, literals, and types, we cannot distinguish between different variables, different types and different method calls anymore. This could render two methods that have an entirely different functionality as clones. Merging such clones can be harmful instead of helpful.

@@

\begin{itemize}
  \item \textbf{Considering types:} Type 2 clones do not consider types. However, this can make a code fragment very hard to refactor, as different types can describe different functional concepts. Because of this, we propose that type 2R clones should consider types like type 1R clones do.
  \item \textbf{Having a distinction between different variables:} For type 2 clones, no identifiers would be taken into account. We agree that a difference in identifiers may still result in a harmful clone, but we should still consider the distinction between different variables. For instance, if we call a method like this: \texttt{myMethod(var1, var2)}, or call this method like this: \texttt{myMethod(var1, var1)}. Even if the variables have the same type, the distinction between the variables is important to ensure the functionality is the same after merging.
  \item \textbf{Defining a threshold for variability in literals:} For type 2 clones no literals would be taken into account. We agree, as when merging the clone (for example by extracting a method), we can turn the literal into a method parameter. However, we would argue that thresholds matter here. How many literals may differ for the segment still to be considered a clone with another segment? We need to define a threshold to be sure that, by merging, we are not replacing a code fragment by a worse maintainable design.
  \item \textbf{Consider method call signatures and define a threshold for variability in method calls:} As type-2 clones allow changes in identifiers, also the names of called methods may vary. However, because of this, completely different methods can be called in cloned fragments as a result. This poses serious challenges on refactoring and makes it more disputable whether such a clone is harmful for the maintainability of the code. This is because different method identifiers can describe a completely different functionality. Therefore, we suggest considering the call signatures of cloned methods when they are compared. We can allow variability in the rest of method identifiers by passing the function as a parameter. To limit the amount of parameters required we also recommend defining a threshold for variability in method call expressions, so only a limited number of method calls can vary.
\end{itemize}

@@

Clone instances in type 3 clones are almost always different in functionality. As we have to ensure equal functionality after merging the clone, we have to wrap the difference in statements between the clone instances in conditional blocks. We can then pass a variable to indicate which path should be taken through the code (either a boolean or an enumeration). Such a refactoring would make added statements that are contiguous less harmful for the design than added statements that are scattered throughout the cloned fragment.

@@

CloneRefactor uses JavaParser \cite{tomassetti2017javaparser}. JavaParser has a built-in symbol solver, which can automatically resolve types, method calls and variables. However, because of this, CloneRefactor does require all libraries that the software project requires (apart from the Java Standard Library). If these are not available, CloneRefactor will estimate types on basis of the information that is available.

@@

\begin{table}[H]
  \begin{center}
  \caption{General results for GitHub Java projects corpus \cite{githubCorpus2013}.} \label{table:general}
  \medskip
\begin{tabular}{|l|l|}
\hline
Amount of projects                                                                                      & 1,361      \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of lines (excluding\\whitespace, comments and newlines.)\end{tabular} & 1,414,996  \\ \hline
Amount of statements/declarations                                                                       & 1,212,189  \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of tokens (excluding\\whitespace, comments and newlines.)\end{tabular} & 11,643,194 \\ \hline
\end{tabular}
\end{center}
\end{table}

@@

\begin{table}[H]
  \begin{center}
  \caption{CloneRefactor clone detection results for the two different algorithms.} \label{table:clonedet}
  \medskip
\begin{tabular}{|l|l|l|}
\hline
 & \textbf{Type 1} & \textbf{Type 1R} \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of lines cloned\end{tabular} & 200,362 & 129,519 \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of statements/\\declarations cloned\end{tabular} & 182,466 & 118,980 \\ \hline
\begin{tabular}[c]{@{}l@{}}Amount of tokens cloned\end{tabular} & 1,582,845 & 973,596 \\ \hline
\end{tabular}
\end{center}
\end{table}
